### 2022-09-30 main/draft.tex
resolve last todos; fix ref [1] noise

This method worked in testing on smeared parton-level simulations from
\madgraph~\cite{Alwall:2014hca} and was presented in two conference posters.
\TODO{Should these posters be included as figures? Maybe appendix}
However, real events have substantial additional noise from QCD
radiation~\cite{PhysRevD.83.074010} and pileup, and this method of approximate,
parametrized modelling does not easily scale to include such extra, complicated
effects.
For this special case it also cost a few second of computing time per event,
which is not competitive with modern Machine Learning algorithms that, once
trained, can be rapidly evaluated over many events and should approximate
the same result.
Although this study shows that we \emph{can} use theoretical modelling
with theoretically motivated algorithms for practical classification, it does
not show that we \emph{should}.
Empirically motivated function approximators such as
Boosted Decision Trees~\cite{xgboost} or
Neural Networks~\cite{MurphyKevinP.2012Mlap} are more scalable to complicated
uses and more computationally efficient in evaluation.

@@

\TODO{\ldots}

@@

% Read2002cls
% }
\TODO{examples of use and misuse in hep}

\TODO{p-value footnote:
The chance of a small \pvalue\ is small.
This is true, but not interesting because the value of any \pvalue\ is small.
By construction \pvalues\ are uniformly distributed on the unit interval
under a given hypothesis.
The frequency of $p < 0.05$ is $5\%$, but so also is the frequency of
$0.6 < p < 0.65$.
The frequency of $p < \varepsilon$ is identical to the frequency
of $0.6 < p < 0.6 + \varepsilon$.
An appropriate choice of test statistic is therefore necessary for the
\pvalue\ to have meaning.
}

@@

\TODO{log maximum likelihood ratio}
$\log$ maximum likelihood ratio $\kappa$
\begin{equation}
\kappa = -\frac{1}{2}\chi^2
\quad
\Leftrightarrow
\quad
\chi = \pm\sqrt{-2\kappa}
\end{equation}

\TODO{tail area}

\begin{align}
P(\chi)
&=
\frac{1}{\sqrt{2\pi}}
\int_\chi^\infty
\!
\exp\!\left(-\frac{1}{2}t^2\right)
\,\mathrm{d}t
\\[0.5em]
&=
L(\chi)
\times
\left(
\frac{1}{\chi}
+
\mathcal{O}\!\left(\frac{1}{\chi^3}\right)
\right)
\end{align}
\TODO{significance measures}
~\cite{
jaynes2003probability,
abramowitz1965handbook
}

@@

\TODO{B-hadrons and LHCb?}
% more cost-effective to improve calorimetry at the expense of PID

@@

\TODO{
Argument:
want to measure W, t,
so measure jets and precise objects
so weaker B and heavy ion programme.
...
Refer to competitive measurements of W and t masses
and xsec of various rare processes (diboson, triboson)
}

@@

To support the later work that this thesis presents on experimental data from
the \atlas\ detector, this chapter introduces contextual information.
 presents the Standard Model of particle
physics, both in its current state and possible future directions.
Section~\ref{sec:context_atlas} then describes the \atlas\ detector.
\TODO{more}

@@

\TODO{This does not belong in an introduction and will be moved.}

\TODO{CUT THIS SHORT}

>>>>

This method worked in testing on smeared parton-level simulations from
\madgraph~\cite{Alwall:2014hca} and was presented in two conference posters.
\NO_DOTO{Should these posters be included as figures? Maybe appendix}
However, real events have substantial additional noise from QCD
radiation~\cite{PhysRevD.83.074010} and pileup, and this method of approximate,
parametrized modelling does not easily scale to include such extra, complicated
effects.
For this special case it also cost a few second of computing time per event,
which is not competitive with modern Machine Learning algorithms that, once
trained, can be rapidly evaluated over many events and should approximate
the same result.
Although this study shows that we \emph{can} use theoretical modelling
with theoretically motivated algorithms for practical classification, it does
not show that we \emph{should}.
Empirically motivated function approximators such as
Boosted Decision Trees~\cite{xgboost} or
Neural Networks~\cite{MurphyKevinP.2012Mlap} are more scalable to complicated
uses and more computationally efficient in evaluation.

@@

\NO_DOTO{\ldots}

@@

% Read2002cls
% }
\NO_DOTO{examples of use and misuse in hep}

\NO_DOTO{p-value footnote:
The chance of a small \pvalue\ is small.
This is true, but not interesting because the value of any \pvalue\ is small.
By construction \pvalues\ are uniformly distributed on the unit interval
under a given hypothesis.
The frequency of $p < 0.05$ is $5\%$, but so also is the frequency of
$0.6 < p < 0.65$.
The frequency of $p < \varepsilon$ is identical to the frequency
of $0.6 < p < 0.6 + \varepsilon$.
An appropriate choice of test statistic is therefore necessary for the
\pvalue\ to have meaning.
}

@@

\NO_DOTO{log maximum likelihood ratio}
$\log$ maximum likelihood ratio $\kappa$
\begin{equation}
\kappa = -\frac{1}{2}\chi^2
\quad
\Leftrightarrow
\quad
\chi = \pm\sqrt{-2\kappa}
\end{equation}

\NO_DOTO{tail area}

\begin{align}
P(\chi)
&=
\frac{1}{\sqrt{2\pi}}
\int_\chi^\infty
\!
\exp\!\left(-\frac{1}{2}t^2\right)
\,\mathrm{d}t
\\[0.5em]
&=
L(\chi)
\times
\left(
\frac{1}{\chi}
+
\mathcal{O}\!\left(\frac{1}{\chi^3}\right)
\right)
\end{align}
\NO_DOTO{significance measures}
~\cite{
jaynes2003probability,
abramowitz1965handbook
}

@@

\NO_DOTO{B-hadrons and LHCb?}
% more cost-effective to improve calorimetry at the expense of PID

@@

\NO_DOTO{
Argument:
want to measure W, t,
so measure jets and precise objects
so weaker B and heavy ion programme.
...
Refer to competitive measurements of W and t masses
and xsec of various rare processes (diboson, triboson)
}

@@

To support the later work that this thesis presents on experimental data from
the \atlas\ detector, this chapter introduces contextual information.
 presents the Standard Model of particle
physics, both in its current state and possible future directions.
Section~\ref{sec:context_atlas} then describes the \atlas\ detector.
\NO_DOTO{more}

@@

\NO_DOTO{This does not belong in an introduction and will be moved.}

\NO_DOTO{CUT THIS SHORT}
